All Graphs Lead to Rome: Learning Geometric and Cycle-Consistent Representations with Graph Convolutional Networks

### 2018-11-17 cvpr2019/main.tex
One more draft - suggestions from Daphne, Davie, Kostas, and Xiao

%%%%%%%%% TITLE
% \title{All Graphs Lead to Rome: \\
% Cycle Consistent Multi-View Representations Using Graph CNNs}
\title{All Graphs Lead to Rome: Learning Geometric and Cycle Consistent Representations with Graph CNNs}

\author{Stephen Phillips, Kostas Daniilidis \\
GRASP Labratory, University of Pennsylvania\\
{\tt\small \{stephi, kostas\}@seas.upenn.edu}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
% \and
% Second Author\\
% Institution2\\
% First line of institution2 address\\
% {\tt\small secondauthor@i2.org}
}

@@

%%%%%%%%% ABSTRACT
% We present a learning technique for refining multi-image feature matches in an unsupervised fashion.
\begin{abstract}
    Image feature matching is a fundametal part of many geometric computer vision applications, and using multiple images can improve performance.
    In this work, we formulate the multi-image matching problem as a graph embedding problem then use a Graph Neural Network to learn an appropriate embedding function for aligning image features. % How to talk about the universe of features here without going too much into the weeds??
    We use cycle consistency to train our network in an unsupervised fashion since ground truth correspondence is difficult or expensive to come by in many real world datasets.
    In addtion, geometric consistency losses can be added at train time, even if the information is not available in the test set, unlike previous optimization based methods.
    To the best of our knowledge, no other works have used learning for multi-image feature matching.
    Our experiments show that our method is competitive with other optimization based approaches.
\end{abstract}

@@

Feature matching is an essential part of Structure from Motion and many geometric computer vision applications.
The goal in multi-image feature matching is to take 2D image locations from three or more images and find which ones correspond to the same point in the 3D scene.
Methods such as SIFT feature matching \cite{lowe2004distinctive} and RANSAC \cite{fischler1981random} have been standard for solving this for decades.
However RANSAC-based approaches are limited for matching pairs of images.
Other works, such as Wang et al. \cite{wang2017multi}, have shown improvement in performance by enforcing consistency between multiple images.
% However we still have not fully answered the question: What are the best image features to describe multiple views of the same scene?

Deep learning has revolutionized how image features are computed \cite{yi2016lift}, and so we would like to leverage this to compute better scene features.
Typically when applied to image or feature matching, deep neural networks (DNNs) are trained with photometric losses, directly using pixel color values.
However direct, pixel-based losses have are not robust to the many confounding factors in the environment, such as lighting and view-point changes.
Using features designed to deal with these condfounding factors improves robustness to them, 
and learning may improve robustness even further.

\begin{figure}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/CycleConsistencyBasic.pdf}
\end{center}
  \caption{
    Example of multi-image matching, with images from Rome16K dataset \cite{li2010location}.
    In this example, an error in matching leads to an inconsistent cycle (shown in red), ideally these pairwise matches would be consistent.
    We improve consistency between pairwise matches in the multi-view matching problem by training a neural net with cycle consistency.
  }
\label{fig:cycconsistex}
\label{fig:onecol}
\end{figure}

%------------------------------------------------------------------------
\begin{figure*}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/CycleConsistencyMainFigure.pdf}
\end{center}
  \caption{
    An outline of the pipeline of this work.
    The Graph Convolutional Neural Network (GCN) \cite{kipf2016semi} takes as input the graph of matches and then outputs a low rank embedding of the adjacency matrix of the graph.
    The GCN operates on an embedding over the nodes of the graph.
    In the figure, the the GCN node embeddings are represented by different colors.
    The final embedding is used to construct a pairwise similarity matrix, which ideally should be a low dimensional cycle consistent representation of the graph adjacency matrix.
    We train the network using a reconstruction loss on the similarity matrix.
    In addition, we can use geometric information at to assist training the network, even if we do not have that geometric information at testing time such as epipolar constraints on the point locations or higher order geometric constraints.
  }
\label{fig:pipeline}
\label{fig:onecol}
\end{figure*}

Unfortunately, there are obstacles to applying multi-view constraints directly to learning. 
To train networks, we need a large amount of labelled data.
In the case of multi-image feature matching, one would need to hand label point correspondences between images, which is difficult and expensive to obtain.
Multi-view constraints are formulated in terms of sparse features, which traditional convolutional deep neural nets (CNNs) are not designed to handle such input.
In the case of multi-image feature matching, geometric constraints would be helpful in rejecting outlier matches.

In this work, we propose to solve these problems using a Graph Neural Network (GNN) to operate on the correspondence graph.
The proposed method works on the correspondence graph, which is agnostic to how the correspondnces were computed, thus allowing the algorithm to work in a broad class of environments.

We use a unsupervised loss, the cycle consistency loss, to train the network.
We add in geometric consistency losses to help training, which works even when the geometric information is not available at test time.
Our contributions are:
\begin{itemize}
\item We use a novel architecture to address multi-image feature matching problem using GNNs with graph embeddings.
\item We introduce an unsupervised cycle consistency loss that does not require labelled correspondences to train.
\item We demonstrate the effectiveness of geometric consistency losses in improving training.
\item We perform experiments on the Rome16K \cite{li2010location} dataset to test the effectiveness of our method compared to optimization based baselines.
\end{itemize}

@@

Matching has a rich history of research in computer vision.
The most well known and widely used method for image matching is RANSAC \cite{fischler1981random}.
RANSAC, combined with hand-crafted feature descriptors like SIFT \cite{lowe2004distinctive}, SURF \cite{bay2006surf}, BRIEF \cite{calonder2012brief}, or ORB \cite{mur2015orb}, has constituted the bulk of the matching literature for the last 40 years.
To improve matching of all the features of the image, graph matching \cite{suh2015subgraph, hu2016distributable} can be used for more robust matches.

Multi-image matching has traditionally been done using optimization based methods minimizing a cycle consistency based loss (see Section 3.3).
Pachauri et al. \cite{pachauri2013solving} and Arrigoni et al. \cite{arrigoni2017synchronization} use the eigenvectors of the matching matrix to obtain a low dimensional embedding. 
Thus they are robust to small Gaussian noise but not to more realistic noise.
Zhou et al. \cite{zhou2015multi} and Wang et al. \cite{wang2017multi} use most sophisticated optimization techniques on the matching matrix and thus are quite robust.
Leonardos et al. \cite{leonardos2016distributed} implement a distributed optimizaiton scheme to solve for cycle consistency.
As an alternative to optimization based techiniques, Tron et al. \cite{tron2017fast} used density based clustering techniques to compute multi-image correspondence.

Previous attempts to improve image matching techniques using machine learning have focused on learning the descriptors given ground truth correspondence from curated datasets \cite{zagoruyko2015learning, yi2016lift, brachmann2017dsac}.
This approach is limited if you do not have the ability to get the ground truth correspondences.
There are other methods to build correspondences such as Choy et al. \cite{choy2016universal}, but it only handles two-view constraints and requires dense correspondences.
Most similar to ours, Yi et al. \cite{yi2018learning} attempts to improve correspondences by learning match probabilites for RANSAC to speed up test time running.
However, they only focus on two view tests and do not exploit the advantages of the correspondence structure.

Graph neural networks have been getting more attention recently \cite{bronstein2017geometric, bruna2013spectral, defferrard2016convolutional, kipf2016semi, scarselli2009graph, gama2018mimo, gama2018convolutional, battaglia2018relational}.
Classical so-called Spectral methods used the eigenvectors of graph Laplacian to compute convolutions like in Bruna et al. \cite{bruna2013spectral}, but requires a fixed graph structure known a-priori. 
Non-spectral methods do not require a-priori knowledge \cite{bronstein2017geometric, kipf2016semi, scarselli2009graph, gama2018convolutional}.
Most of these simply use polynomials of the graph Laplacian to compute neighborhood averages. Gama et al. 2018 \cite{gama2018mimo, gama2018convolutional} formalize this notion and generalize it beyond the use of the graph Laplacian.
To improve performance, more sophisticated aggregation techniques and global information passing can be used as discussed in Battaglia et al. \cite{battaglia2018relational}.

@@

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.9\linewidth]{figures/UniverseOfFeatures.pdf}
\end{center}
  \caption{
    Illustration of the universe of features.
    Each feature in each image corresponds to a 3D point in the scene.
    Therefore, we can get the points to have a cycle consistent embedding by mapping each feature in each image to the one-hot vector of its corresponding 3D point.
    While there can be many features, there are much fewer 3D points and thus this corrsponds to a low rank factorization of the corrspondence matrix.
    Best viewed in color.
  }
\label{fig:universefeatures}
\label{fig:onecol}
\end{figure}
Our goal is to learn optimal features that capture multiple image views by way of filtering out noisy feature matches.
The input to our algorithm is a set of features and noisy correspondences, and the output is a new set of features where the pairwise similarities of these features correspond to the true matches.
We do this by learning training the new set of feature embeddings to be cycle consistent.
We formulate this problem in terms of the correspondence graph of the features.
Vectors $\mat{x}$ and matrices $\mat{A}$ are denoted with boldface, and the $i^{th}$ row of a matrix is denoted $(\mat{A})_i$.
\subsection{Correspondence Graph}
We assume we have an initial set of feature matches represented as a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$.
Each vertex $v$ of the graph is a feature with its associated descriptor $f_v$. 
The graph is represented by its Weighted Adjacency Matrix $\mat{A}(\mathcal{G}) \in \bR^{n \times n}$, where $(\mat{A}(\mathcal{G}))_{ij}$ is the strength of the match between nodes $i$ and $j$. We also have the positive diagonal degree matrix $\mat{D}(\mathcal{G}) \in \bR^{n \times n}$, with $\mat{D}(\mathcal{G})_{ii} = \sum_j (\mat{A}(\mathcal{G}))_{ij}$. For brevity we denote these matrices $\mat{A}$ and $\mat{D}$ respectively.
We can get our an embedding matrix $\mat{E}_0$ by concatenating the features $f_v$ from each of the vertices.
We use $\mat{E}_0$ as the initializiation for our learning algorithm.
We add any additional knowledge we have into the embedding as well (e.g. scale, orientation, etc.).
Putative correspondences are computed from these features and represented by weighted edges in which the weight gives the strength of the match.
While there are many interesting methods for computing these putative correspondences \cite{suh2015subgraph, yi2018learning}, we do not explore them in this work.

In the absense of noise or outliers, this graph would have a connected component for each visible point in the world.
These components should be mutually disjoint. 
In other words each feature would only have edges to other features corresponding to the same point.
Since features in this case represent unique location in the scene, no points in the same image would have edges between them.
In the noisy case we expect this structure to be corrupted and thus we need to refine out the erroneous edges.

To compute this matching graph, we do pairwise feature matching between the images, creating putative correspondences for each of the features.
Typically these putative correspondences are matched probabilistically, meaning a feature in one image matches to many features in another.
This could come from repeated structures in the scene, ambiguity from the low-level feature descriptors, or just an error in the matching algorithm.
Filtering out these noisy matches is our primary learning goal.

However, this graph structure does not have a regular grid structure and thus we cannot use standard CNNs to learn features for this task.
Instead we use graph convolutions to learn feature representations on this space.
We describe our approach in the next section.

\subsection{Graph Neural Networks}
As input to our method we are given the graph Laplacian $\mat{L} \in \bR^{n \times n}$ to encode the graph structure:
\begin{equation}
  \mat{L} =\; \mat{I} - \mat{D}^{-1/2} \mat{A} \mat{D}^{-1/2} \\
\end{equation}
and an initial embedding $\mat{E}_0 \in \bR^{n \times m_0}$.
Many older methods assume $\mat{L}$ is known a-priori \cite{bruna2013spectral}, and thus can encode graph convolutions using the eigenvectors of $L$.
We do not have this luxury, as the correspondence structure changes from image to image, and thus we use non-spectral GNNs. The standard form of non-spectral GNNs is:
\begin{equation}
\mat{E}_{i+1} =\; P\left(\sigma\left( \sum_k \mat{L}^k \mat{E}_i \mat{W}_{i,k} \right)\right)
\end{equation}
Here $\sigma$ is the non-linearity (typically a ReLU), $\mat{W}_{k,i}$ are the learned weights of layer $i$,  and $P$ is the `pooling' operation, known as graph coarsening \cite{bronstein2017geometric, gama2018mimo}.
The maximum degree of $\mat{L}$ determines the number of hops away from a node one layer can access.
In the feature matching problem, we require an embedding for each vertex.
Therefore we cannot perform graph coarsening, and thus we only perform the non-linearity step.
We instead use a modification of the approach proposed in Kipf and Welling \cite{kipf2016semi}:
\begin{align}
      \widetilde{\mat{L}} =&\; (\mat{D} + \mat{I})^{-1/2} (\mat{A} + \mat{I}) (\mat{D} + \mat{I})^{-1/2} \\
\mat{E}_{l+1} =&\; \sigma\left(\widetilde{\mat{L}} \mat{E}_i \mat{W}_i \right)  \label{eq:graph_conv}
\end{align}

The matrix $\widetilde{\mat{L}}$ is analogous to the graph Laplacian (with better numerical stability properties).
$\widetilde{\mat{L}}$ encodes the structure of the graph and is used to perform the actual graph convolution.
Our network has many layers, and thus the receptive field of the final embedding is quite large, even without the pooling operations.
In principle we could use higher hop neighborhoods \cite{gama2018convolutional}, or more complicated aggregation structures \cite{battaglia2018relational}, but in this work we restrict ourselves to simpler architectures, which still shows promising results for this task.
For faster training, we use group normalization \cite{wu2018group} and skip connections.
The final output $\mat{E}_n$ gives us a new descriptor for every node.
In the next sections we will discuss the loss functions to train this network.

@@

Let $\mat{M}$ be the noiseless set of matches between our features, with $\mat{M}_{ij}$ being the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then we can say that, for all $i, j, k$:
\begin{equation}
\mat{M}_{ij} = \mat{M}_{ik} \mat{M}_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stays the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle concistency}, and has been used in a number of works to optimize for global consistency \cite{zhou2015multi, wang2017multi, leonardos2016distributed}.
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a matrix $\mat{X}_i$.
Then the cycle consistency becomes:
\begin{equation}
\mat{M}_{ij} = \mat{X}_{i}\mat{X}_{j}^\top
\label{eq:cycconsist2}
\end{equation}

This reduces our complexity from $O(n^3)$ to $O(n^2)$.
We try to learn $\mat{E}$ to approximate $\mat{X}$ - in other words the final embedding should be an encoding of the universe of features. As we do not have the ground truth matches $\mat{M}$, we approximate it using the noisy adjacency matrix $\mat{A}$ of our correspondence graph. Thus our loss would be 
\begin{equation}
\mathcal{L}(\mat{A}, \mat{E}_n) = \mathcal{D}(\mat{A}, \mat{E}_n\mat{E}_n^\top)
\end{equation}
Here $\mathcal{D}$ could be an $L_2$ loss, $L_1$ loss, or many others. In this work, we use the $L_1$ loss.

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.8\linewidth]{figures/GeometricConsistency.pdf}
\end{center}
  \caption{
    Illustrated here is an example of how the geometric loss on one feature.  
    We compute the errors on each feature match to determine their validity.
    Errors are computed via absolute distance from the epipolar line, as expressed by (\ref{eq:essential_constraint}) via the epipolar constraint.
    Best viewed in color.
  }
\label{fig:geoconsist}
\label{fig:onecol}
\end{figure}

@@

\begin{figure*}
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/ExampleOutput.pdf}
  \end{center}
     \caption{Example output of our network. (a) Similarity Matrix of the of the embeddings (b) Histogram of feature similarities for pairs which correspond to the same 3D point, comparing our output features with the original SIFT features (c) Histogram of feature similarities for pairs which correspond to different 3D points, comparing our output features with the original SIFT features}
  \label{fig:short}
\end{figure*}
\subsection{Synthetic Graph Dataset}
We first test our method on synthetically generated data as a simple proof of concept.
To generate the data, we generate $p$ points, each with its own randomly generated descriptor.
To create the graph, we generate random permutation matrices, with a noise applied to it after generated.
The initial descriptors we use as input we created using the true descriptor plus some added Gaussian noise.
No geometric losses were added during training for these experiments.
However, the method was robust in testing with different noise functions and parameters.
The normalized noisy input descriptors are our baseline - they correlate with the true values but do not perserve the structure well.
However, with the GNN we are able to recover the true structure very well, as shown in \ref{fig:synthtable}.
All experiments were run with a 12 layer GCN with the ReLU nonlinearity and skip connections.
All were trained with the Adam optimizer \cite{kingma2014adam} and a learning rate $10^{-4}$

>>>>

%%%%%%%%% TITLE
% \title{All Graphs Lead to Rome: \\
% Cycle Consistent Multi-View Representations Using Graph CNNs}
\title{All Graphs Lead to Rome: Learning Geometric and Cycle-Consistent Representations with Graph Convolutional Networks}

\author{Stephen Phillips, Kostas Daniilidis \\
GRASP Laboratory, University of Pennsylvania\\
{\tt\small \{stephi, kostas\}@seas.upenn.edu}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
% \and
% Second Author\\
% Institution2\\
% First line of institution2 address\\
% {\tt\small secondauthor@i2.org}
}

@@

%%%%%%%%% ABSTRACT
% We present a learning technique for refining multi-image feature matches in an unsupervised fashion.
\begin{abstract}
    Image feature matching is a fundamental part of many geometric computer vision applications, and using multiple images can improve performance.
    In this work, we formulate multi-image matching as a graph embedding problem then use a Graph Convolutional Network to learn an appropriate embedding function for aligning image features.
    We use cycle consistency to train our network in an unsupervised fashion, since ground truth correspondence is difficult or expensive to aquire.
    In addition, geometric consistency losses can be added at train time, even if the information is not available in the test set, unlike previous approaches that optimize cycle consistency directly.
    To the best of our knowledge, no other works have used learning for multi-image feature matching.
    Our experiments show that our method is competitive with other optimization based approaches.
\end{abstract}

@@

Feature matching is an essential part of Structure from Motion and many geometric computer vision applications.
The goal in multi-image feature matching is to take 2D image locations from three or more images and find which ones correspond to the same point in the 3D scene.
Methods such as SIFT feature matching \cite{lowe2004distinctive} followed by RANSAC \cite{fischler1981random} have been the standard for decades.
However RANSAC-based approaches are limited for matching pairs of images, which can lead to global inconsistencies in the matching.
Other works, such as Wang et al.~\cite{wang2017multi}, have shown improvement in performance by optimizing cycle consistency, i.e. enforcing the pairwise feature matches to be globally consistent, illustrated in \ref{fig:cycconsistex}.

However, these multi-view consistency algorithms struggle in distributed and robust settings.
Deep learning has revolutionized how image features are computed \cite{yi2016lift}.
In this paper, we want to leverage the power of deep representations in order to compute feature descriptors that are robust across multiple views.
% Typically when applied to image or feature matching, deep neural networks (DNNs) are trained with photometric losses, directly using pixel color values.
% However direct, pixel-based losses have are not robust to the many confounding factors in the environment, such as lighting and view-point changes.
% Using features designed to deal with these confounding factors improves robustness to them, and learning may improve robustness even further.

\begin{figure}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/CycleConsistencyBasic.pdf}
\end{center}
  \caption{
    Example of multi-image matching and cycle consistency, with images from Rome16K dataset \cite{li2010location}.
    Pairwise matches between image can lead to global inconsistencies in the matching.
    In this example, an error in matching leads to an inconsistent cycle (shown in red), ideally these pairwise matches would be consistent.
    We improve the quality of pairwise matches in the multi-view matching problem by training a neural net with cycle consistency.
  }
\label{fig:cycconsistex}
\label{fig:onecol}
\end{figure}

%------------------------------------------------------------------------
\begin{figure*}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/CycleConsistencyMainFigure.pdf}
\end{center}
  \caption{
    An illustration of the approach of this work.
    The Graph Convolutional Neural Network (GCN) \cite{kipf2016semi} takes as input the graph of matches and then outputs a low rank embedding of the adjacency matrix of the graph.
    The GCN operates on an embedding over the nodes of the graph.
    In the figure, the GCN node embeddings are represented by different colors.
    The final embedding is used to construct a pairwise similarity matrix, which we train to be a low dimensional cycle-consistent representation of the graph adjacency matrix, thus pruning the erroneous matches.
    We train the network using a reconstruction loss on the similarity matrix with the noisy adjacency matrix, and thus do not need ground truth matches.
    In addition, we can use geometric consistency information, such as epipolar constraints, to assist training the network.
  }
\label{fig:pipeline}
\label{fig:onecol}
\end{figure*}

Unfortunately, there are obstacles to applying multi-view constraints directly to deep learning. 
To train networks, we need a large amount of labeled data.
In the case of multi-image feature matching, one would need to hand label point correspondences between images, which is difficult and expensive to obtain.
Multi-view constraints are formulated in terms of sparse features, which traditional convolutional neural nets are not designed to handle.
In the case of multi-image feature matching, geometric constraints would be helpful in rejecting outlier matches.

In this work, we propose to solve these problems using a Graph Convolutional Network (GCN) to operate on the correspondence graph.
The proposed method works directly on the correspondence graph, which is agnostic to how the correspondences were computed, thus allowing the algorithm to work in a broad class of environments.
To the best of our knowledge this work is the first to apply deep learning to the multi-view feature matching problem.
We use an unsupervised loss - the cycle consistency loss - to train the network, and thus avoiding the difficulty of expensive hand labeling.
% We incorporate geometric consistency losses to guide training, which helps training even when the geometric information is not available at test time.
Geometric consistency losses can aid training, even if such information is not available at test time.
Although our network is simple, it shows promising results compared to baselines which optimize for cycle-consistency without learned embeddings.
Furthermore, since inference requires only a single forward pass over the neural network, our approach is faster (to achieve comparable accuracy) than method which must solve an optimization problem every time.
Our contributions are:
\begin{itemize}
\item We use a novel architecture to address the multi-image feature matching problem using GCNs with graph embeddings.
\item We introduce an unsupervised cycle consistency loss that does not require labeled correspondences to train.
\item We demonstrate the effectiveness of geometric consistency losses in improving training.
\item We perform experiments on the Rome16K \cite{li2010location} dataset to test the effectiveness of our method compared to optimization based methods.
\end{itemize}

@@

\subsection{Feature Matching}
Matching has a rich history of research in computer vision.
The most well known and widely used method for image matching is RANSAC \cite{fischler1981random}.
RANSAC, combined with hand-crafted feature descriptors such as SIFT \cite{lowe2004distinctive}, SURF \cite{bay2006surf}, BRIEF \cite{calonder2012brief}, or ORB \cite{mur2015orb}, has constituted the bulk of the matching literature for the last 40 years.
Finally, graph matching \cite{suh2015subgraph, hu2016distributable} can be used as a final step for more robust matches.

\subsection{Multi-image Matching}
Multi-image matching has traditionally been done using optimization based methods minimizing a cycle consistency based loss (see Section 3.3).
Pachauri et al.~\cite{pachauri2013solving} and Arrigoni et al.~\cite{arrigoni2017synchronization} use the eigenvectors of the matching matrix to obtain a low dimensional embedding. 
However, the low Gaussian noise assumption is not realistic.
Zhou et al.~\cite{zhou2015multi} and Wang et al.~\cite{wang2017multi} use most sophisticated optimization techniques on the matching matrix and thus produce more robust solutions.
Leonardos et al.~\cite{leonardos2016distributed} implement a distributed optimization scheme to solve for cycle consistency.
As an alternative to optimization based techniques, Tron et al.~\cite{tron2017fast} used density based clustering techniques to compute multi-image correspondence.
To the best of our knowledge, we are the first to use neural networks for multi-image matching.

\subsection{Deep Learning for Matching}
Previous attempts to improve image matching techniques using machine learning have focused on learning the descriptors given ground truth correspondence from curated datasets \cite{zagoruyko2015learning, yi2016lift, brachmann2017dsac}.
This approach is limited if one do not have the ability to get the ground truth correspondences.
There are other methods to build correspondences such as Choy et al.~\cite{choy2016universal}, but they only handle two-view constraints and require dense correspondences.
Most similar to our work, Yi et al.~\cite{yi2018learning} attempts to improve correspondences by learning match probabilities for RANSAC for greater robustness and speed.
However, they only focus on two view matching and do not exploit the advantages of the correspondence structure.
Note that while Zhu et al.~\cite{zhu2017unpaired} use cycle consistency in their loss, their method is restricted to pairwise cycle consistency, while we use multi-image cycle consistency.

\subsection{Graph Neural Networks}
Graph neural networks have recieved more attention recently \cite{bronstein2017geometric, bruna2013spectral, defferrard2016convolutional, kipf2016semi, scarselli2009graph, gama2018mimo, gama2018convolutional, battaglia2018relational}.
Classical so-called Spectral methods used the eigenvectors of the graph Laplacian to compute convolutions as in Bruna et al.~\cite{bruna2013spectral}, but requires an a-prior known graph structure. 
Non-spectral methods do not require a-priori knowledge \cite{bronstein2017geometric, kipf2016semi, scarselli2009graph, gama2018convolutional}.
Most of these methods use polynomials of the graph Laplacian to compute neighborhood averages.
Gama et al.~2018 \cite{gama2018mimo, gama2018convolutional} formalize this notion and generalize it beyond the use of the graph Laplacian.
To improve performance, more sophisticated aggregation techniques and global information passing can be used as discussed in Battaglia et al.~\cite{battaglia2018relational}.

@@

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.9\linewidth]{figures/UniverseOfFeatures.pdf}
\end{center}
  \caption{
    Illustration of the universe of features.
    Each feature in each image corresponds to a 3D point in the scene.
    We can construct a cycle consistent embeddings of the features by mapping each one to the one-hot vector of its corresponding 3D point.
    While there can be many features, there are fewer 3D points and thus this corresponds to a low rank factorization of the correspondence matrix.
    Best viewed in color.
  }
\label{fig:universefeatures}
\label{fig:onecol}
\end{figure}
Our goal is to learn optimal features that capture multiple image views by filtering out noisy feature matches.
The input to our algorithm is a set of features and noisy correspondences, and the output is a new set of features where the pairwise similarities of these features correspond to the true matches.
We do this by training the new set of feature embeddings to be cycle consistent.
We formulate this problem in terms of the correspondence graph of the features.
Vectors $\mat{x}$ and matrices $\mat{A}$ are denoted with boldface, and the $i^{th}$ row of a matrix is denoted $(\mat{A})_i$.
\subsection{Correspondence Graph}
We assume there is an initial set of feature matches represented as a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$.
Each vertex $v$ of the graph is a feature with its associated descriptor $f_v$. 
The graph is represented by its Weighted Adjacency Matrix $\mat{A}(\mathcal{G}) \in \bR^{n \times n}$, where $(\mat{A}(\mathcal{G}))_{ij}$ is the strength of the match between nodes $i$ and $j$.
We introduce the positive diagonal degree matrix $\mat{D}(\mathcal{G}) \in \bR^{n \times n}$, with $\mat{D}(\mathcal{G})_{ii} = \sum_j (\mat{A}(\mathcal{G}))_{ij}$.
For brevity we denote these matrices $\mat{A}$ and $\mat{D}$ respectively.
We introduce the notion of an embedding matrix $\mat{E}_0$ by concatenating the features $f_v$ from each of the vertices.
We use $\mat{E}_0$ as the initialization for our learning algorithm.
We add any additional knowledge we have into the embedding as well (e.g. scale, orientation, etc.).
Putative correspondences are computed from these features and represented by weighted edges in which the weight gives the strength of the match.
While there are many interesting methods for computing these putative correspondences \cite{suh2015subgraph, yi2018learning}, we do not explore them in this work.

In the absence of noise or outliers, this graph would have a connected component for each visible point in the world.
These components should be mutually disjoint. 
In other words each feature would only have edges to other features corresponding to the same point.
Since features in this case represent unique locations in the scene, no points in the same image would have edges between them.
In the noisy case we expect this structure to be corrupted and thus we need to prune the erroneous edges.

To compute this matching graph, we do pairwise feature matching between the images, creating putative correspondences for each of the features.
Typically these putative correspondences are matched probabilistically, meaning a feature in one image matches to many features in another.
The ambiguity in the matches could come from repeated structures in the scene, insufficiently informative low-level feature descriptors, or just an error in the matching algorithm.
Filtering out these noisy matches is our primary learning goal.

However, this graph structure does not have a regular grid structure and thus we cannot use standard convolutional nets to learn features for this task.
Instead we use graph convolutions to learn feature representations on this space.
We describe our approach in the next section.

\subsection{Graph Neural Networks}
As input to our method we are given the graph Laplacian $\mat{L} \in \bR^{n \times n}$ to encode the graph structure:
\begin{equation}
  \mat{L} =\; \mat{I} - \mat{D}^{-1/2} \mat{A} \mat{D}^{-1/2} \\
\end{equation}
and an initial embedding $\mat{E}_0 \in \bR^{n \times m_0}$.
Many older methods assume $\mat{L}$ is known a-priori \cite{bruna2013spectral}, and can encode graph convolutions using the eigenvectors of $L$.
We do not have this luxury, as the correspondence structure changes from image to image, and thus we use non-spectral Graph Neural Networks.
The standard form of non-spectral Graph Neural Networks is:
\begin{equation}
\mat{E}_{i+1} =\; P\left(\sigma\left( \sum_k \mat{L}^k \mat{E}_i \mat{W}_{i,k} \right)\right)
\end{equation}
Here $\sigma$ is the non-linearity (typically a ReLU), $\mat{W}_{k,i}$ are the learned weights of layer $i$,  and $P$ is the `pooling' operation, known as graph coarsening \cite{bronstein2017geometric, gama2018mimo}.
The maximum degree of $\mat{L}$ determines the number of hops away from a node one layer can access.
In the feature matching problem, we require an embedding for each vertex.
Therefore we cannot perform graph coarsening, and thus only perform the non-linearity step.
This work uses Graph Convolutional Networks, proposed in Kipf and Welling \cite{kipf2016semi}:
\begin{align}
      \widetilde{\mat{L}} =&\; (\mat{D} + \mat{I})^{-1/2} (\mat{A} + \mat{I}) (\mat{D} + \mat{I})^{-1/2} \\
\mat{E}_{l+1} =&\; \sigma\left(\widetilde{\mat{L}} \mat{E}_i \mat{W}_i \right)  \label{eq:graph_conv}
\end{align}

The matrix $\widetilde{\mat{L}}$ is analogous to the graph Laplacian (with better numerical stability properties).
$\widetilde{\mat{L}}$ encodes the structure of the graph and is used to perform the actual graph convolution.
Our network has many layers, and thus the receptive field of the final embedding is quite large, even without the pooling operations.
In principle we could use higher hop neighborhoods \cite{gama2018convolutional}, or more complicated aggregation structures \cite{battaglia2018relational}, but in this work we restrict ourselves to simpler architectures, and still obtain promising results for this task.
For faster training, we use group normalization \cite{wu2018group} and skip connections.
The final output $\mat{E}_n$ gives us a new descriptor for every node.
In the next sections we will discuss the loss functions to train this network.

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.8\linewidth]{figures/GeometricConsistency.pdf}
\end{center}
  \caption{
    Illustrated here is an example of how the geometric loss is computed for one feature.  
    Errors are computed via absolute distance from the epipolar line, as expressed by (\ref{eq:essential_constraint}) via the epipolar constraint.
    The epipolar line is the line of projection of the feature on the first image, projected onto to the second.
    The distance to this line on the second image indicates how likely that point is to correspond geometrically to the original feature.
    There can be false positives along the projected line, as shown by the square feature in the figure, but other points will be eliminated, such as the hexagonal feature.
    Best viewed in color.
  }
\label{fig:geoconsist}
\label{fig:onecol}
\end{figure}

@@

Let $\mat{M}$ be the noiseless set of matches between our features, with $\mat{M}_{ij}$ being the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then we can say that, for all $i, j, k$:
\begin{equation}
\mat{M}_{ij} = \mat{M}_{ik} \mat{M}_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stay the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle consistency}, and has been used in a number of works to optimize for global consistency \cite{zhou2015multi, wang2017multi, leonardos2016distributed}.
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a matrix $\mat{X}_i$.
Then the cycle consistency becomes:
\begin{equation}
\mat{M}_{ij} = \mat{X}_{i}\mat{X}_{j}^\top
\label{eq:cycconsist2}
\end{equation}

@@

\begin{figure*}
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures/ExampleOutput.pdf}
  \end{center}
     \caption{Example output of our network. (a) Similarity Matrix of the of the embeddings (b) Histogram of feature similarities for pairs which correspond to the same 3D point, comparing our output features with the original SIFT features (c) Histogram of feature similarities for pairs which correspond to different 3D points, comparing our output features with the original SIFT features}
  \label{fig:short}
\end{figure*}
\subsection{Synthetic Graph Dataset}
We first test our method on synthetically generated data as a simple proof of concept.
To generate the data, we generate $p$ points, each with its own randomly generated descriptor.
To create the graph, we generate random permutation matrices, with a noise applied to it after generated.
% The initial descriptors we use as input we created using the true descriptor plus some added Gaussian noise.
We initialize the input descriptors using the true descriptor, plus some added Gaussian noise.
No geometric losses were added during training for these experiments.
However, the method was robust in testing with different noise functions and parameters.
The normalized noisy input descriptors are our baseline - they correlate with the true values but do not preserve the structure well.
However, the GCN recovered the true structure very well, as shown in \ref{fig:synthtable}.
All experiments were run with a 12 layer GCN with the ReLU nonlinearity and skip connections.
All were trained with the Adam optimizer \cite{kingma2014adam} and a learning rate $10^{-4}$
